---
title: "Data Programming Project 2024"
author: "Lucas Lorenzo Jakin"
data: "2024-06-04"
format: html
output: html_document
editor: visual
---

# Predicting Secondary School Student Performance Using Machine Learning

## Student Performance Dataset

Education is a key factor for achieving a long-term economic progress. During the last decades, the portuguese educational level has improved. However, the statistics keep Portugal at Europe's tail end due to its *high student failure* and *dropping rates*.

In particular, failure in the core classes of Mathematics and Portuguese (native language) is extremely serious, since they provide fundamental knowledge for the success in other remaining subjects.

The representing [kaggle dataset](https://www.kaggle.com/datasets/devansodariya/student-performance-data/data) contains real-world data from two Portuguese secondary schools. Data was collected using two sources: mark reports, and questionnaires.

------------------------------------------------------------------------

```{r}
library(tidyverse)
library(dplyr)
library(skimr)
library(gt)
dataset <- read.csv("student_data.csv", header = T)
```

Now about the dataframe; The initial version contained scarce information (only the grades and number of absences were available), it was complemented with the latter, which allowed the collection of several demographic, social and school related attributes. The final dataframe consists of 33 columns (attributes), where G3 represents the target variable, the one that we want to predict.

> The dataset looks like the following:

```{r}
  head(dataset)
  skim(dataset)
```

The dataset information above is shown by using the **skim()** function from the **skimr package**, it provides summary statistics about variables in data frames.

As we can see from the Data Summary, the attribute values are distributed into two data types: String (Character) and Numerical data.

------------------------------------------------------------------------

\

@tbl-attributes listed below shows all the columns that are present inside the data frame combined with their descriptions and data type.

```{r}
library(knitr)
att <- colnames(dataset)
description <- c("student's school(binary)","student's sex(binary)" ,"student's age(Num)","home address","family size(binary)" ,"parent's cohabition status(binary)","mother's education(Num)","father's education(Num)","mother's job(nominal)","father's job(nominal)","reason to choose school(nominal)","student's guardian(nominal)","travel time (Num)","weekly study time(Num)","past class failures(Num)","extra school support(binary)","family educational support(binary)","extra paid classes(binary)","extra-curricular activities(binary)","attended nursery school(binary)","wants to take higher education(binary)","internet access(binary)","romantic relationship(binary)","quality of family relationship(Num)","free time(Num)","going out w/ friends(Num)","workday alcohol consumption(Num)","weekend alcohol consumption(Num)","health status(Num)","absences (Num)","first period grade(Num)","second period grade(Num)","final grade(Num)")

table<- data.frame(Attribute = att, Description = description)


```

```{r}
#| label: tbl-attributes
#| tbl-cap: "The preprocessed student related variables"
  table %>% gt()

```

------------------------------------------------------------------------

# Exploratary Data Analysis (EDA)

Classification and regression are two important DM goals. Both require a supervised learning, but they are different in terms of the output representation(descrete for classification and continuous for regression). In this work, the Mathematics and Poruguese grades (i.e. G3 of @tbl-attributes) will be modeled using three supervised approches:

-   **Binary** Classification: *pass* if G3 \geq 10, else *fail;*

-   **5-Level** Classification: based on *Erasmus* grade conversion system;

-   **Regression** of the G3 value : Numeric output between 0 and 20.

***

Now let's look at key insights and patterns within the dataset. The analysis focuses on the examination of the attributes distribution , identifying relationships between the attributes and visualizing the data through scatter plots, histograms and other graphical methods. The findigs will then provide a robust foundation for any subsequent analysis or modeling efforts.
  
## Overall Health vs Grade
```{r}
dataset %>% 
```

